---
title: "noaa-sentiments"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(sentimentr)
library(jsonlite)
```
## Look at sentiment of NOAA rulese since 2001.
Goal is to see if 2020 rules post-COVID are more positive than pre-COVID.

## Get data from the federal register
Pull from the federal register
Currently a hard-coded search for NOAA from 2001-01-01 to present, type = rules

```{r pulldata}
#temp$count
#[1] 600137

# xx <- fromJSON("https://www.federalregister.gov/api/v1/documents?conditions%5Bagencies%5D%5B%3Dnational-oceanic-and-atmospheric-administration%5D=&conditions%5Bpublication_date%5D%5Bgte%5D%5B%3D2001-01-01%5D=&format=json&page=0",flatten = TRUE)

pull_regs <- function(x) {
page <- paste0("https://www.federalregister.gov/api/v1/documents?conditions%5Bagencies%5D%5B%5D=national-oceanic-and-atmospheric-administration&conditions%5Bpublication_date%5D%5Bgte%5D=01%2F01%2F2001&conditions%5Bpublication_date%5D%5Blte%5D=05%2F28%2F2020&conditions%5Btype%5D%5B%5D=RULE&format=json&page=",x,"&per_page=1000")
 xx <- fromJSON(page, flatten = TRUE)$results
 return(xx)
}
#get number of pages in search
page <- paste0("https://www.federalregister.gov/api/v1/documents?conditions%5Bagencies%5D%5B%5D=national-oceanic-and-atmospheric-administration&conditions%5Bpublication_date%5D%5Bgte%5D=01%2F01%2F2001&conditions%5Bpublication_date%5D%5Blte%5D=05%2F28%2F2020&conditions%5Btype%5D%5B%5D=RULE&format=json&page=0&per_page=1000")
fpage <- fromJSON(page, flatten = TRUE)
npage <- fpage$total_pages

plan(multiprocess)
system.time(regs <- future_map_dfr(1:npage,pull_regs))
range(regs$publication_date)
regs <- tibble(regs)
```

## Look at the sentiment of the abstracts of these rules

Using `sentimentr` package. to find average sentiment of sentences in the abstracts.    

```{r sentiment}
# rules by NOAA
regs_part <- regs %>%
   unnest(cols = "agencies") %>% 
   filter(name == "National Oceanic and Atmospheric Administration",
          type == "Rule") %>%
  I()
regs_part

# sentiments of sentences of the abstract and then average over abstracts
regs_sentiment <- regs_part %>% 
  select(publication_date, title, type, abstract, document_number) %>% 
  rowid_to_column() %>% 
  mutate(sentences = map(abstract,get_sentences),
         sentiment = map(sentences,sentiment_by)) %>% 
  unnest(cols = sentiment) %>% 
  I()


# make publication date more easy to work with
regs_summary <- regs_sentiment %>% 
  select(publication_date, ave_sentiment) %>% 
  mutate(date = lubridate::date(publication_date),
         year = lubridate::year(publication_date)) %>% 
  I()

```

## viz

summarize by year

```{r plot-sentiment}


regs_summary %>%
  ggplot() +
  #aes(y = ave_sentiment, x = factor(year)) +
  aes(y = ave_sentiment, x = date) +
  #geom_boxplot() +
  geom_jitter(alpha = 0.5) +
  geom_smooth() +
  NULL

regs_summary %>%
  ggplot() +
  aes(y = ave_sentiment, x = factor(year)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "year",
       y = "average sentiment",
       title = "NOAA rules in federal register more positive in 2020",
       subtitle = "average sentiment of abstracts of NOAA Rules",
       caption = "\n@gavin_fay\ndata from https://federalregister.gov") +
  #geom_jitter(alpha = 0.5) +
  #geom_smooth() +
  NULL

ggsave("noaa-sentiment-yr.png",width = 8, height = 4.5)
```


```{r}
regs_summary %>%
  mutate(month = lubridate::month(date)) %>% 
  filter(month <= 5) %>% 
  ggplot() +
  aes(y = ave_sentiment, x = factor(year)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "year",
       y = "average sentiment",
       title = "NOAA rules in federal register more positive in 2020",
       subtitle = "average sentiment of abstracts of NOAA Rules",
       caption = "\n@gavin_fay\ndata from https://federalregister.gov") +
  #geom_jitter(alpha = 0.5) +
  #geom_smooth() +
  NULL

```

A qnd linear model suggests positive effect of 2020 is statistically clear.

```{r}
regs_lm <- regs_summary %>%   
  mutate(month = lubridate::month(date)) %>% 
  filter(month <= 5) %>% 
  mutate(twenty = ifelse(year==2020, 1, 0)) %>% 
  #lm(ave_sentiment~twenty,data=.) %>% 
  lm(ave_sentiment~factor(year)-1,data=.) %>%   
  tidy(conf.int = TRUE) %>% #augment()
  I()
regs_lm
regs_lm %>% 
  mutate(year = 2001:2020) %>% 
  ggplot() +
  aes(x=year, y = estimate) +
  #geom_jitter(data = regs_summary, aes(x=year, y=ave_sentiment), alpha = 0.2) +
  geom_point() +
  geom_errorbar(aes(ymin=conf.low, ymax = conf.high)) +
  geom_smooth(data = regs_summary, mapping = aes(x=year, y = ave_sentiment), method = "lm") +
  geom_smooth(data = regs_summary, mapping = aes(x=year, y = ave_sentiment), method = "gam", col = "orange") +
  #ylim(-0.1,0.4) +
  labs(x = "year",
       y = "average sentiment",
       title = "NOAA rules in federal register more positive in 2020",
       subtitle = "average sentiment of abstracts of NOAA Rules during Jan-May",
       caption = "\n@gavin_fay\ndata from https://federalregister.gov") +
  NULL
```

```{r}
tt <- " NMFS approves and
implements the New England Fishery
Management Council’s Habitat Clam
Dredge Exemption Framework
Adjustment to its Fishery Management
Plans. This action establishes three
areas within the Great South Channel
Habitat Management Area where vessels
may not fish for Atlantic surfclams or blue
mussels with dredge gear. This action is
intended to restrict the fishing industry
access to part of the surfclam and blue
mussel resource within the Habitat
Management Area while balancing the
Council’s habitat conservation
objectives. "
sentiment_by(tt)
```

