---
title: "Changes in sentiment of US federal fisheries regulations during COVID-19"
author:
  - Gavin Fay:
      email: gfay@umassd.edu
      institute: [SMAST]
      correspondence: true
  - Scott I. Large:
      email: fl@another.edu
      institute: [NEFSC]
      correspondence: false
institute:
  - SMAST: Department of Fisheries Oceanography, School for Marine Science and Technology, University of Massachusetts Dartmouth, New Bedford, MA 02744, USA
  - NEFSC: NOAA Fisheries Northeast Fisheries Science Center, Woods Hole, MA 02543, USA
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::pdf_document2:
      fig_caption: yes
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
 The COVID-19 crisis has had a significant impact on the US seafood industry, with disruption to both fishing activity and distribution of products to markets, particularly with restaurant closures. Reductions in fishing activity have been concomitant with efforts by fishers and supporting organizations to pivot to alternative markets and new initiatives to sell and distribute products. In addition, both state and federal fisheries regulators and management agencies have made changes to rules that affect fishing behavior, including monitoring requirements and shifts in governing when, where, and what fishers can catch. A strategic goal of fisheries management in the US is to reduce burdens of regulatory complexity on fishers, as a means of promoting increased production from wild capture fisheries. Here we perform sentiment analysis of US federal fisheries rules published in the Code of Federal Regulations to assess whether recent rule changes during the COVID-19 crisis have used language suggestive of relaxing constraints on fishers (e.g. easing regulations), and compare changes in rules sentiment to those observed during other recent large-scale shocks to the US seafood system. Finally, we evaluate changes in US fisheries regulation texts in the context of changes within the entire US Federal Register system.
keywords: |
  COVID-19; management complexity; sentiment
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The following code chunk defines some general settings how code chunks should behave. -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)
```

<!-- The actual document text starts here: -->

# Introduction

Here is a citation [@Marwick2017]
The COVID-19 crisis has had a significant impact on the seafood industry, with closures and restrictions to both fishing activity and distribution of seafood products to markets creating a large shock to the seafood supply chain. These impacts affect both fishers and also other businesses that are part of the seafood supply chain in and around working waterfronts and coastal communities.

In concert and in response to these changes, US fisheries have seen both large scale and regional changes to regulations, including designation of CARES act funding, chiefly aimed at providing relief to fishers, to reduce restrictions on fishing activity, and also to promote workforce safety, notably the waiver of observer requirements because of risks associated with COVID-19 spread.
	Text here supporting claims about changes in regulations

As NOAA Fisheries moves toward implementing Integrated Ecosystem Approaches to management, require systemic indicators across both strategic and tactical objectives.
Indicators for governance system
Strategic plans (and EO) calls for reducing regulatory burdens on fisheries
How to quantify these?

Here we perform sentiment analysis of US federal fisheries rules published in the code of federal regulations to assess whether recent rule changes during the COVID-19 crisis have used language suggestive of relaxing constraints on fishers (e.g. easing regulations), and compare changes in rules sentiment to those observed during other recent large-scale shocks to the US seafood system (implementation of catch shares). Finally, we evaluate changes in US fisheries regulation texts in the context of changes within the entire US federal registry system

# Methods

## Data 

Table \@ref(tab:cfr-part-table))

* Subpart (e.g., Groundfish plan)
* Body to evaluate
    + FR - actions, summary of regulations (abstract), most recent action, # of pages 
    + CFR - (eCFR)
* Implementation of catch shares created a more restrictive/ volume set of rules
* Sentiment analysis - trends in #s
* What words are correlating with sentiment score (reduce, waive, relax, extension)
    + Think about what exactly we want to measure.
* # pages as proxy of complexity, sentiment analysis on summary & on whole regulation
* Compare effect relative to catch shares (or ACLs) 
* Resampling distributions



```{r cfr-part-table}
part_df <- data.frame(Part = c(600, 622, 635, 648, 660, 665, 679, 680, 697),
                      Name = c("Magnuson-Stevens Act Provisions",
                               "Fisheries of the Caribbean, Gulf of Mexico, and South Atlantic",
                               "Atlantic Highly Migratory Species", 
                               "Fisheries of the Northeastern United States",
                               "Fisheries Off West Coast States",
                               "Fisheries in the Western Pacific",
                               "Fisheries of the Exclusive Economic Zone Off Alaska",
                               "Shellfish Fisheries of the Exclusive Economic Zone Off Alaska",
                               "Atlantic Coastal Fisheries Cooperative Management"))

knitr::kable(part_df, caption = "Title 50---Wildlife and Fisheries; CHAPTER VI--Fishery Conservation and Management, National Oceanic and Atmospheric Administration, Department of Commerce")

```
<br>


# Results

* General trend in #s of regulations, #of pages.
* Sentiment of rules in 2020 compared to other years
    + Split by part & subpart
    + Linear model to aid interpretation
* Comparison to catch share, at same resolution
* Comparison to resample distribution for random parts & subparts


<!-- Here's some example analysis code: -->

```{r get-data, eval = FALSE}
# Note the path that we need to use to access our data files when rendering this document
my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
```

```{r demo-plot, fig.cap="A plot of random numbers"}
plot(rnorm(10))
```

Figure \@ref(fig:demo-plot) shows how we can have a caption and cross-reference for a plot

```{r demo-inline-code}
x <- round(pi, 2)
```

Here is an example of inline code `r x` in the middle of a sentence. 

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break  -->

\newpage

# References 

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

<div id="refs"></div>

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies: 

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```
